---
title: "ELAN Dashboard - Basic Municipality Analysis Report"
author: "ELAN Data Center"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    highlight: tango
params:
  n: NA
  selected_variable: NA
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Load required libraries
library(dplyr)
library(ggplot2)
library(knitr)
library(kableExtra)
library(scales)
library(tidyr)

# Custom CSS for consistent styling
cat('
<style>
body {
  font-family: Arial, Tahoma, Geneva, Verdana, sans-serif;
  line-height: 1.6;
  color: #333;
}
h1, h2, h3, h4, h5, h6 {
  color: #0e1d6b;
  font-weight: 600;
}
.table {
  font-size: 14px;
}
.alert {
  padding: 15px;
  margin-bottom: 20px;
  border: 1px solid transparent;
  border-radius: 4px;
}
.alert-info {
  color: #31708f;
  background-color: #d9edf7;
  border-color: #bce8f1;
}
.alert-success {
  color: #3c763d;
  background-color: #dff0d8;
  border-color: #d6e9c6;
}
.alert-warning {
  color: #8a6d3b;
  background-color: #fcf8e3;
  border-color: #faebcc;
}
</style>
')

# Initialize variables
selected_municipalities <- if(exists("params") && !is.null(params$n)) {
  params$n
} else {
  character(0)
}

# Function to get latest year for specific variables
get_latest_year_for_variables <- function(data, variables) {
  if(nrow(data) == 0) return(0)
  if(inherits(data, "sf")) {
    data <- as.data.frame(data)
  }
  existing_variables <- variables[variables %in% names(data)]
  if(length(existing_variables) == 0) return(0)
  filtered_data <- data %>%
    filter(if_any(all_of(existing_variables), ~!is.na(.)))
  if(nrow(filtered_data) == 0) return(0)
  return(max(filtered_data$YEAR, na.rm = TRUE))
}

# Get selected municipalities data with year-specific filtering
if (length(selected_municipalities) > 0 && exists("states_gmc")) {
  states_gmc_df <- as.data.frame(states_gmc)
  demographic_vars <- c("Total_Population", "Total_ICPCPat_Pop", "AGE_MEAN", "Income_MEAN", 
                        "Age0to20", "Age21to40", "Age41to60", "Age61to80", "AgeAbove80")
  healthcare_cost_vars <- c("ZVWKOSTENTOTAAL_MEAN", "ZVWKZIEKENHUIS_MEAN", "ZVWKHUISARTS_MEAN", "ZVWKFARMACIE_MEAN")
  primary_care_vars <- c("ZVWKHUISARTS_user")
  medication_vars <- c("UniqueMed_Count5")
  latest_demographic_year <- get_latest_year_for_variables(states_gmc_df, demographic_vars)
  latest_healthcare_cost_year <- get_latest_year_for_variables(states_gmc_df, healthcare_cost_vars)
  latest_primary_care_year <- get_latest_year_for_variables(states_gmc_df, primary_care_vars)
  latest_medication_year <- get_latest_year_for_variables(states_gmc_df, medication_vars)
  demographic_data <- states_gmc_df %>%
    filter(GMN %in% selected_municipalities, YEAR == latest_demographic_year) %>%
    select(GMN, YEAR, all_of(demographic_vars)) %>%
    distinct(GMN, .keep_all = TRUE)
  if(nrow(demographic_data) == 0) {
    demographic_data <- data.frame(GMN = selected_municipalities, YEAR = latest_demographic_year)
  }
  healthcare_cost_data <- states_gmc_df %>%
    filter(GMN %in% selected_municipalities, YEAR == latest_healthcare_cost_year) %>%
    select(GMN, all_of(healthcare_cost_vars)) %>%
    distinct(GMN, .keep_all = TRUE)
  if(nrow(healthcare_cost_data) == 0) {
    healthcare_cost_data <- data.frame(GMN = selected_municipalities)
  }
  primary_care_data <- states_gmc_df %>%
    filter(GMN %in% selected_municipalities, YEAR == latest_primary_care_year) %>%
    select(GMN, all_of(primary_care_vars)) %>%
    distinct(GMN, .keep_all = TRUE)
  if(nrow(primary_care_data) == 0) {
    primary_care_data <- data.frame(GMN = selected_municipalities)
  }
  medication_data <- states_gmc_df %>%
    filter(GMN %in% selected_municipalities, YEAR == latest_medication_year) %>%
    select(GMN, all_of(medication_vars)) %>%
    distinct(GMN, .keep_all = TRUE)
  if(nrow(medication_data) == 0) {
    medication_data <- data.frame(GMN = selected_municipalities)
  }
  selected_data <- demographic_data %>%
    left_join(healthcare_cost_data, by = "GMN") %>%
    left_join(primary_care_data, by = "GMN") %>%
    left_join(medication_data, by = "GMN")
  time_series_data <- states_gmc_df %>%
    filter(GMN %in% selected_municipalities) %>%
    distinct(GMN, YEAR, .keep_all = TRUE) %>%
    arrange(YEAR)
} else {
  if(exists("states_gmc")) {
    states_gmc_df <- as.data.frame(states_gmc)
    sample_municipalities <- unique(states_gmc_df$GMN)[1:3]
    demographic_vars <- c("Total_Population", "Total_ICPCPat_Pop", "AGE_MEAN", "Income_MEAN", 
                          "Age0to20", "Age21to40", "Age41to60", "Age61to80", "AgeAbove80")
    healthcare_cost_vars <- c("ZVWKOSTENTOTAAL_MEAN", "ZVWKZIEKENHUIS_MEAN", "ZVWKHUISARTS_MEAN", "ZVWKFARMACIE_MEAN")
    primary_care_vars <- c("ZVWKHUISARTS_user")
    medication_vars <- c("UniqueMed_Count5")
    latest_demographic_year <- get_latest_year_for_variables(states_gmc_df, demographic_vars)
    latest_healthcare_cost_year <- get_latest_year_for_variables(states_gmc_df, healthcare_cost_vars)
    latest_primary_care_year <- get_latest_year_for_variables(states_gmc_df, primary_care_vars)
    latest_medication_year <- get_latest_year_for_variables(states_gmc_df, medication_vars)
    demographic_data <- states_gmc_df %>%
      filter(GMN %in% sample_municipalities, YEAR == latest_demographic_year) %>%
      select(GMN, YEAR, all_of(demographic_vars)) %>%
      distinct(GMN, .keep_all = TRUE)
    if(nrow(demographic_data) == 0) {
      demographic_data <- data.frame(GMN = sample_municipalities, YEAR = latest_demographic_year)
    }
    healthcare_cost_data <- states_gmc_df %>%
      filter(GMN %in% sample_municipalities, YEAR == latest_healthcare_cost_year) %>%
      select(GMN, all_of(healthcare_cost_vars)) %>%
      distinct(GMN, .keep_all = TRUE)
    if(nrow(healthcare_cost_data) == 0) {
      healthcare_cost_data <- data.frame(GMN = sample_municipalities)
    }
    primary_care_data <- states_gmc_df %>%
      filter(GMN %in% sample_municipalities, YEAR == latest_primary_care_year) %>%
      select(GMN, all_of(primary_care_vars)) %>%
      distinct(GMN, .keep_all = TRUE)
    if(nrow(primary_care_data) == 0) {
      primary_care_data <- data.frame(GMN = sample_municipalities)
    }
    medication_data <- states_gmc_df %>%
      filter(GMN %in% sample_municipalities, YEAR == latest_medication_year) %>%
      select(GMN, all_of(medication_vars)) %>%
      distinct(GMN, .keep_all = TRUE)
    if(nrow(medication_data) == 0) {
      medication_data <- data.frame(GMN = sample_municipalities)
    }
    selected_data <- demographic_data %>%
      left_join(healthcare_cost_data, by = "GMN") %>%
      left_join(primary_care_data, by = "GMN") %>%
      left_join(medication_data, by = "GMN")
    time_series_data <- states_gmc_df %>%
      filter(GMN %in% sample_municipalities) %>%
      distinct(GMN, YEAR, .keep_all = TRUE) %>%
      arrange(YEAR)
  } else {
    selected_data <- data.frame()
    time_series_data <- data.frame()
  }
}

# Store latest years for use in titles
latest_years <- list(
  demographic = if(exists("latest_demographic_year")) latest_demographic_year else 0,
  healthcare_cost = if(exists("latest_healthcare_cost_year")) latest_healthcare_cost_year else 0,
  primary_care = if(exists("latest_primary_care_year")) latest_primary_care_year else 0,
  medication = if(exists("latest_medication_year")) latest_medication_year else 0
)

# Calculate summary statistics
if (nrow(selected_data) > 0) {
  cost_vars_available <- c("ZVWKOSTENTOTAAL_MEAN", "ZVWKZIEKENHUIS_MEAN", "ZVWKHUISARTS_MEAN", "ZVWKFARMACIE_MEAN")
  cost_vars_available <- cost_vars_available[cost_vars_available %in% names(selected_data)]
  summary_stats <- selected_data %>%
    summarise(
      total_population = sum(Total_Population, na.rm = TRUE),
      total_elan_patients = sum(Total_ICPCPat_Pop * Total_Population, na.rm = TRUE),
      avg_age = mean(AGE_MEAN, na.rm = TRUE),
      avg_income = mean(Income_MEAN, na.rm = TRUE),
      avg_total_healthcare_cost = if("ZVWKOSTENTOTAAL_MEAN" %in% names(selected_data)) mean(ZVWKOSTENTOTAAL_MEAN, na.rm = TRUE) else 0,
      avg_secondary_care_cost = if("ZVWKZIEKENHUIS_MEAN" %in% names(selected_data)) mean(ZVWKZIEKENHUIS_MEAN, na.rm = TRUE) else 0,
      avg_primary_care_cost = if("ZVWKHUISARTS_MEAN" %in% names(selected_data)) mean(ZVWKHUISARTS_MEAN, na.rm = TRUE) else 0,
      primary_care_users = sum(ZVWKHUISARTS_user * Total_Population, na.rm = TRUE),
      primary_care_non_users = sum(Total_Population - (ZVWKHUISARTS_user * Total_Population), na.rm = TRUE),
      poly_pharmacy_users = sum(UniqueMed_Count5 * Total_Population, na.rm = TRUE),
      avg_medication_cost = if("ZVWKFARMACIE_MEAN" %in% names(selected_data)) mean(ZVWKFARMACIE_MEAN, na.rm = TRUE) else 0
    )
  summary_stats <- summary_stats %>%
    mutate(across(everything(), ~ifelse(is.na(.), 0, .)))
} else {
  summary_stats <- data.frame(
    total_population = 0,
    total_elan_patients = 0,
    avg_age = 0,
    avg_income = 0,
    avg_total_healthcare_cost = 0,
    avg_secondary_care_cost = 0,
    avg_primary_care_cost = 0,
    primary_care_users = 0,
    primary_care_non_users = 0,
    poly_pharmacy_users = 0,
    avg_medication_cost = 0
  )
}
```

# ELAN Dashboard - Basic Municipality Analysis Report

<div style="text-align: center; margin-bottom: 30px; padding: 20px; background: linear-gradient(135deg, #f6fafe 0%, #e8f4fd 100%); border-radius: 10px;">
  <h1 style="color: #0e1d6b; margin-bottom: 10px; font-size: 32px; font-weight: bold;">ELAN Dashboard</h1>
  <h2 style="color: #007CC2; margin-bottom: 10px; font-size: 24px;">Basic Municipality Analysis Report</h2>
  <p style="color: #666; font-size: 16px; margin: 0;">Health Campus Den Haag</p>
  <p style="color: #666; font-size: 14px; margin-top: 10px;">Generated on `r Sys.Date()`</p>
</div>

<div style="background-color: #f8f9fa; padding: 20px; border-radius: 8px; margin-bottom: 30px; border-left: 4px solid #007CC2;">
  <h2 style="color: #0e1d6b; margin-bottom: 15px;">📊 Executive Summary</h2>
  <p style="color: #333; line-height: 1.6; margin-bottom: 10px;">
    This basic analysis examines the selected municipalities focusing on key healthcare and demographic indicators. The report provides insights into population characteristics and healthcare utilization patterns using the most recent available data for each indicator type.
  </p>
  <p style="color: #333; line-height: 1.6;">
    <strong>Data Coverage:</strong><br>
    • Demographics and Population: `r if(latest_years$demographic > 0) latest_years$demographic else "N/A"`<br>
    • Healthcare Costs: `r if(latest_years$healthcare_cost > 0) latest_years$healthcare_cost else "N/A"`<br>
    • Primary Care Utilization: `r if(latest_years$primary_care > 0) latest_years$primary_care else "N/A"`<br>
    • Medication Usage: `r if(latest_years$medication > 0) latest_years$medication else "N/A"`<br>
    <strong>Municipalities Analyzed:</strong> `r if(length(unique(selected_data$GMN)) > 0) paste(unique(selected_data$GMN), collapse = ", ") else "No municipalities selected"`
  </p>
  `r if(is.null(params$n) || length(params$n) == 0) '<div style="background-color: #fff3cd; padding: 15px; border-radius: 5px; margin-top: 15px; border-left: 4px solid #ffc107;"><p style="color: #856404; margin: 0;"><strong>Note:</strong> No specific municipalities were selected. This report shows sample data from the first 3 available municipalities for demonstration purposes.</p></div>' else ''`
</div>

## Geographic Overview

```{r choropleth-map, echo=FALSE, fig.width=12, fig.height=8, warning=FALSE}
# Load required libraries for mapping
library(ggplot2)
library(sf)
library(dplyr)
library(RColorBrewer)
library(scales)

# Create static choropleth map function
create_static_choropleth_map <- function(spatial_data, selected_areas, variable_name = "Total_ICPCPat_Pop", area_type = "gemeente") {
  
  # Check if we have spatial data
  if (!exists("spatial_data") || nrow(spatial_data) == 0) {
    # Create a simple placeholder plot
    plot(1, 1, type = "n", axes = FALSE, xlab = "", ylab = "")
    text(1, 1, "No spatial data available", cex = 1.5, col = "gray")
    return()
  }
  
  # Convert to data frame if it's sf object
  if (inherits(spatial_data, "sf")) {
    spatial_df <- as.data.frame(spatial_data)
  } else {
    spatial_df <- spatial_data
  }
  
  # Get the latest year with data
  if ("YEAR" %in% names(spatial_df)) {
    latest_year <- max(spatial_df$YEAR, na.rm = TRUE)
    map_data <- spatial_df %>% filter(YEAR == latest_year)
  } else {
    map_data <- spatial_df
  }
  
  # If no selected areas, use all available areas
  if (length(selected_areas) == 0) {
    if (area_type == "wijk") {
      selected_areas <- unique(map_data$WKN)[1:min(3, length(unique(map_data$WKN)))]
    } else {
      selected_areas <- unique(map_data$GMN)[1:min(3, length(unique(map_data$GMN)))]
    }
  }
  
  # Check if variable exists
  if (!variable_name %in% names(map_data)) {
    variable_name <- "Total_ICPCPat_Pop"  # Default fallback
  }
  
  # Filter data for selected areas
  if (area_type == "wijk") {
    filtered_data <- map_data %>% filter(WKN %in% selected_areas)
    name_col <- "WKN"
    muni_col <- "GMN"
  } else {
    filtered_data <- map_data %>% filter(GMN %in% selected_areas)
    name_col <- "GMN"
    muni_col <- "GMN"
  }
  
  # If no data after filtering, create placeholder
  if (nrow(filtered_data) == 0) {
    plot(1, 1, type = "n", axes = FALSE, xlab = "", ylab = "")
    text(1, 1, "No data available for selected areas", cex = 1.5, col = "gray")
    return()
  }
  
  # Get spatial data for plotting
  if (area_type == "wijk" && exists("states_wkc")) {
    plot_data <- states_wkc %>% filter(YEAR == latest_year)
    plot_data$selected <- plot_data$WKN %in% selected_areas
    plot_data$highlight <- plot_data$WKN %in% selected_areas
  } else if (area_type == "gemeente" && exists("states_gmc")) {
    plot_data <- states_gmc %>% filter(YEAR == latest_year)
    plot_data$selected <- plot_data$GMN %in% selected_areas
    plot_data$highlight <- plot_data$GMN %in% selected_areas
  } else {
    # Create placeholder if no spatial data
    plot(1, 1, type = "n", axes = FALSE, xlab = "", ylab = "")
    text(1, 1, "No spatial data available", cex = 1.5, col = "gray")
    return()
  }
  
  # Create color palette based on variable values
  var_values <- plot_data[[variable_name]]
  var_values <- var_values[!is.na(var_values)]
  
  if (length(var_values) == 0) {
    # If no valid values, use a simple color scheme
    plot_data$color_value <- 1
    color_palette <- "Blues"
  } else {
    # Create color scale based on variable values
    plot_data$color_value <- plot_data[[variable_name]]
    color_palette <- "Blues"
  }
  
  # Get the proper variable label for the legend
  if (exists("var_def_label_dict") && variable_name %in% names(var_def_label_dict)) {
    variable_label <- var_def_label_dict[[variable_name]]
  } else {
    variable_label <- variable_name
  }
  
  # Create the map
  p <- ggplot() +
    # Base map - all areas with variable values
    geom_sf(data = plot_data, 
            aes(fill = color_value),
            color = "#666666", 
            size = 0.5) +
    # Highlight selected areas with yellow border
    geom_sf(data = plot_data %>% filter(highlight), 
            fill = "yellow", 
            color = "#ff8f00", 
            size = 2) +
    # Color scale for variable values with improved legend
    scale_fill_gradientn(
      colors = brewer.pal(9, color_palette),
      name = variable_label,
      labels = comma,
      na.value = "#f0f0f0"
    ) +
    # Theme settings with bigger legend
    theme_minimal() +
    theme(
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      panel.grid = element_blank(),
      plot.title = element_text(size = 18, face = "bold", color = "#0e1d6b"),
      plot.subtitle = element_text(size = 14, color = "#666"),
      legend.position = "bottom",
      legend.title = element_text(size = 14, face = "bold"),
      legend.text = element_text(size = 12),
      legend.key.height = unit(1.5, "cm"),
      legend.key.width = unit(2, "cm"),
      legend.margin = margin(t = 10, b = 10, l = 10, r = 10)
    ) +
    labs(
      title = paste("Geographic Distribution -", variable_label),
      subtitle = paste("Selected areas highlighted in yellow | Year:", latest_year),
      x = NULL,
      y = NULL
    )
  
  return(p)
}

# Determine which variable to show on the map
# Use the selected variable from the dashboard control panel
if (exists("params") && !is.null(params$selected_variable)) {
  # Use the variable selected in the dashboard
  map_variable <- params$selected_variable
} else if (nrow(selected_data) > 0) {
  # Try to find a good variable to display from available data
  potential_vars <- c("Total_ICPCPat_Pop", "AGE_MEAN", "Income_MEAN", 
                      "ZVWKOSTENTOTAAL_MEAN", "ZVWKZIEKENHUIS_MEAN", "ZVWKHUISARTS_MEAN",
                      "UniqueMed_Count5", "Gender_Mannen", "Gender_Vrouwen")
  
  map_variable <- NULL
  for (var in potential_vars) {
    if (var %in% names(selected_data) && any(!is.na(selected_data[[var]]))) {
      map_variable <- var
      break
    }
  }
  
  if (is.null(map_variable)) {
    map_variable <- "Total_ICPCPat_Pop"  # Better default fallback
  }
} else {
  map_variable <- "Total_ICPCPat_Pop"
}

# Get the proper variable label for the legend
if (exists("var_def_label_dict") && map_variable %in% names(var_def_label_dict)) {
  variable_label <- var_def_label_dict[[map_variable]]
} else {
  variable_label <- map_variable
}

# Create the map
if (exists("states_gmc") && length(selected_municipalities) > 0) {
  # Use selected municipalities
  selected_areas <- selected_municipalities
} else if (exists("states_gmc")) {
  # Use sample municipalities
  states_gmc_df <- as.data.frame(states_gmc)
  selected_areas <- unique(states_gmc_df$GMN)[1:min(3, length(unique(states_gmc_df$GMN)))]
} else {
  selected_areas <- character(0)
}

# Create the choropleth map with improved legend
choropleth_map <- create_static_choropleth_map(states_gmc, selected_areas, map_variable, "gemeente")
print(choropleth_map)
```

<div style="background-color: #fff3cd; padding: 15px; border-radius: 5px; margin: 20px 0; border-left: 4px solid #ffc107;">
  <p style="color: #856404; margin: 0;">
    <strong>Map Legend:</strong> The map above shows the geographic distribution of the analyzed municipalities with color-coded variable values. 
    Selected municipalities are highlighted with a <strong>yellow border</strong>, while the color intensity represents the variable values. 
    Darker colors indicate higher values, lighter colors indicate lower values.
  </p>
</div>

## Key Findings

<div style="background-color: #e8f4fd; padding: 20px; border-radius: 8px; margin: 20px 0; border-left: 4px solid #007CC2;">
  <h3 style="color: #0e1d6b; margin-bottom: 15px;">🎯 Key Findings</h3>
  <ul style="color: #333; line-height: 1.6;">
    <li><strong>Population Overview:</strong> `r if(summary_stats$total_population > 0) paste("Total population of", format(summary_stats$total_population, big.mark=","), "residents across", length(unique(selected_data$GMN)), "municipalities") else "Population data unavailable"`</li>
    <li><strong>ELAN Patients:</strong> `r if(summary_stats$total_elan_patients > 0) paste("Total ELAN patients:", format(round(summary_stats$total_elan_patients), big.mark=","), "(", round(summary_stats$total_elan_patients / summary_stats$total_population * 100, 1), "% of population)") else "ELAN patient data unavailable"`</li>
    <li><strong>Healthcare Costs:</strong> `r if(summary_stats$avg_total_healthcare_cost > 0) paste("Average total healthcare cost: €", format(round(summary_stats$avg_total_healthcare_cost), big.mark=",")) else "Healthcare cost data unavailable"`</li>
    <li><strong>Primary Care:</strong> `r if(summary_stats$primary_care_users > 0) paste("Primary care users:", format(round(summary_stats$primary_care_users), big.mark=","), "(", round(summary_stats$primary_care_users / summary_stats$total_population * 100, 1), "% of population)") else "Primary care data unavailable"`</li>
    <li><strong>Primary Care:</strong> `r if(summary_stats$primary_care_non_users > 0) paste("Non Primary care users:", format(round(summary_stats$primary_care_non_users), big.mark=","), "(", round(summary_stats$primary_care_non_users / summary_stats$total_population * 100, 1), "% of population)") else "Primary care data unavailable"`</li>
    <li><strong>Polypharmacy:</strong> `r if(summary_stats$poly_pharmacy_users > 0) paste("Polypharmacy users (≥5 meds):", format(round(summary_stats$poly_pharmacy_users), big.mark=","), "(", round(summary_stats$poly_pharmacy_users / summary_stats$total_population * 100, 1), "% of population)") else "Polypharmacy data unavailable"`</li>
    <li><strong>Demographic Profile:</strong> `r if(summary_stats$avg_age > 0) paste("Average age of", round(summary_stats$avg_age, 1), "years") else "Age data unavailable"`</li>
    <li><strong>Economic Status:</strong> `r if(summary_stats$avg_income > 0) paste("Average income of €", format(round(summary_stats$avg_income), big.mark=",")) else "Income data unavailable"`</li>
  </ul>
</div>

---

## Detailed Analysis

### 1. ELAN Patient Trends

```{r elan-trends, echo=FALSE, fig.width=10, fig.height=6}
# Create ELAN patient trends
if (nrow(time_series_data) > 0) {
  required_cols <- c("GMN", "YEAR", "Total_ICPCPat_Pop", "Total_Population")
  missing_cols <- setdiff(required_cols, names(time_series_data))
  if (length(missing_cols) == 0) {
    trend_data <- as.data.frame(time_series_data) %>%
      select(GMN, YEAR, Total_ICPCPat_Pop, Total_Population) %>%
      mutate(
        Total_ICPCPat_Pop = ifelse(is.na(Total_ICPCPat_Pop), 0, Total_ICPCPat_Pop),
        Total_Population = ifelse(is.na(Total_Population), 0, Total_Population),
        ELAN_Patients = Total_ICPCPat_Pop * Total_Population
      )
    if (nrow(trend_data) > 0) {
      elan_plot <- trend_data %>%
        ggplot(aes(x = YEAR, y = ELAN_Patients, color = GMN)) +
        geom_line(size = 1) +
        geom_point(size = 2) +
        labs(
          title = "ELAN Patient Trends",
          subtitle = "Number of ELAN Patients Over Time",
          x = "Year",
          y = "ELAN Patients",
          color = "Municipality"
        ) +
        theme_minimal() +
        theme(
          plot.title = element_text(size = 18, face = "bold", color = "#0e1d6b"),
          plot.subtitle = element_text(size = 14, color = "#666"),
          legend.position = "bottom"
        ) +
        scale_color_brewer(palette = "Set1") +
        scale_y_continuous(labels = comma)
      print(elan_plot)
    } else {
      cat("No valid trend data available for plotting.")
    }
  } else {
    cat("Missing required columns for trend analysis:", paste(missing_cols, collapse = ", "))
  }
} else {
  cat("No time series data available for trend analysis.")
}
```

```{r elan-table, echo=FALSE}
# Create ELAN patient summary table
if (nrow(selected_data) > 0) {
  required_cols <- c("GMN", "Total_Population", "Total_ICPCPat_Pop", "AGE_MEAN", "Income_MEAN")
  missing_cols <- setdiff(required_cols, names(selected_data))
  if (length(missing_cols) == 0) {
    elan_summary_table <- as.data.frame(selected_data) %>%
      select(GMN, Total_Population, Total_ICPCPat_Pop, AGE_MEAN, Income_MEAN) %>%
      mutate(
        Total_Population = ifelse(is.na(Total_Population), 0, Total_Population),
        Total_ICPCPat_Pop = ifelse(is.na(Total_ICPCPat_Pop), 0, Total_ICPCPat_Pop),
        AGE_MEAN = ifelse(is.na(AGE_MEAN), 0, AGE_MEAN),
        Income_MEAN = ifelse(is.na(Income_MEAN), 0, Income_MEAN),
        ELAN_Patients = Total_ICPCPat_Pop * Total_Population,
        ELAN_Rate = ifelse(Total_Population > 0, Total_ICPCPat_Pop, 0)
      ) %>%
      arrange(desc(ELAN_Rate)) %>%
      select(GMN, Total_Population, ELAN_Patients, ELAN_Rate, AGE_MEAN, Income_MEAN)
    if (nrow(elan_summary_table) > 0 && ncol(elan_summary_table) == 6) {
      elan_summary_table_formatted <- elan_summary_table %>%
        mutate(
          Total_Population = format(Total_Population, big.mark = ",", scientific = FALSE),
          ELAN_Patients = format(round(ELAN_Patients, 0), big.mark = ",", scientific = FALSE),
          ELAN_Rate = sprintf("%.2f", ELAN_Rate *100),
          AGE_MEAN = sprintf("%.2f", AGE_MEAN),
          Income_MEAN = format(round(Income_MEAN, 0), big.mark = ",", scientific = FALSE)
        )
      kable(elan_summary_table_formatted,
            col.names = c("Municipality", "Total Population", "ELAN Patients", "ELAN Rate (%)", "Average Age", "Average Income (€)"),
            caption = paste("ELAN Patient Summary by Municipality", if(latest_years$demographic > 0) paste("(", latest_years$demographic, ")") else "")) %>%
        kable_styling(bootstrap_options = c("striped", "hover"),
                      full_width = FALSE,
                      position = "center") %>%
        column_spec(1, bold = TRUE, color = "#0e1d6b") %>%
        row_spec(0, bold = TRUE, color = "white", background = "#007CC2")
    } else {
      cat("Unable to create ELAN summary table due to data structure issues.")
    }
  } else {
    cat("Missing required columns for ELAN summary table:", paste(missing_cols, collapse = ", "))
  }
} else {
  cat("No data available for ELAN summary table.")
}
```

**Key Trends:**
- **ELAN Patients:** `r if(nrow(time_series_data) > 0 && !all(is.na(time_series_data$Total_ICPCPat_Pop))) paste("ELAN patients", if(max(time_series_data$Total_ICPCPat_Pop * time_series_data$Total_Population * time_series_data$Total_Population / 100, na.rm=TRUE) > min(time_series_data$Total_ICPCPat_Pop * time_series_data$Total_Population * time_series_data$Total_Population / 100, na.rm=TRUE)) "increased" else "decreased", "from", format(round(min(time_series_data$Total_ICPCPat_Pop * time_series_data$Total_Population * time_series_data$Total_Population / 100, na.rm=TRUE)), big.mark=","), "to", format(round(max(time_series_data$Total_ICPCPat_Pop * time_series_data$Total_Population * time_series_data$Total_Population / 100, na.rm=TRUE)), big.mark=","), "patients") else "No ELAN patient trend data available"`
- **Peak Utilization:** `r if(nrow(time_series_data) > 0 && !all(is.na(time_series_data$Total_ICPCPat_Pop * time_series_data$Total_Population))) paste("Highest ELAN patient count observed in", time_series_data$YEAR[which.max(time_series_data$Total_ICPCPat_Pop * time_series_data$Total_Population * time_series_data$Total_Population / 100)]) else "No utilization data available"`

### 2. Age Distribution Analysis

```{r age-distribution, echo=FALSE, fig.width=10, fig.height=6}
# Create age distribution analysis
if (nrow(selected_data) > 0) {
  age_vars <- c("Age0to20", "Age21to40", "Age41to60", "Age61to80", "AgeAbove80")
  age_vars_available <- age_vars[age_vars %in% names(selected_data)]
  if (length(age_vars_available) > 0) {
    age_data <- as.data.frame(selected_data) %>%
      select(GMN, all_of(age_vars_available)) %>%
      mutate(
        across(all_of(age_vars_available), ~ifelse(is.na(.), 0, . * 100))
      ) %>%
      gather(key = "Age_Group", value = "Population", -GMN) %>%
      mutate(
        Age_Group = factor(Age_Group,
                          levels = c("Age0to20", "Age21to40", "Age41to60", "Age61to80", "AgeAbove80"),
                          labels = c("0-20", "21-40", "41-60", "61-80", "80+"))
      )
    if (nrow(age_data) > 0) {
      age_plot <- age_data %>%
        ggplot(aes(x = Age_Group, y = Population, fill = GMN)) +
        geom_bar(stat = "identity", position = "dodge") +
        labs(
          title = "Age Distribution Across Municipalities",
          subtitle = "Population by Age Groups",
          x = "Age Group",
          y = "Population (%)",
          fill = "Municipality"
        ) +
        theme_minimal() +
        theme(
          plot.title = element_text(size = 18, face = "bold", color = "#0e1d6b"),
          plot.subtitle = element_text(size = 14, color = "#666"),
          legend.position = "bottom"
        ) +
        scale_fill_brewer(palette = "Set2") +
        scale_y_continuous(labels = comma)
      print(age_plot)
    } else {
      cat("No valid age distribution data available for plotting.")
    }
  } else {
    cat("No age category variables available in the dataset.")
  }
} else {
  cat("No data available for age distribution analysis.")
}
```

```{r age-table, echo=FALSE}
# Create age distribution summary table
if (nrow(selected_data) > 0) {
  age_vars <- c("Age0to20", "Age21to40", "Age41to60", "Age61to80", "AgeAbove80")
  age_vars_available <- age_vars[age_vars %in% names(selected_data)]
  if (length(age_vars_available) > 0) {
    age_summary_table <- as.data.frame(selected_data) %>%
      select(GMN, Total_Population, all_of(age_vars_available)) %>%
      mutate(
        across(all_of(age_vars_available), ~ifelse(is.na(.), 0, . * 100))
      ) %>%
      arrange(desc(Age0to20))
    if (nrow(age_summary_table) > 0) {
      age_summary_table_formatted <- age_summary_table %>%
        mutate(
          across(all_of(age_vars_available), ~sprintf("%.2f", .))
        )
      col_names <- c("Municipality", "Total Population")
      if("Age0to20" %in% age_vars_available) col_names <- c(col_names, "Age 0-20 (%)")
      if("Age21to40" %in% age_vars_available) col_names <- c(col_names, "Age 21-40 (%)")
      if("Age41to60" %in% age_vars_available) col_names <- c(col_names, "Age 41-60 (%)")
      if("Age61to80" %in% age_vars_available) col_names <- c(col_names, "Age 61-80 (%)")
      if("AgeAbove80" %in% age_vars_available) col_names <- c(col_names, "Age 80+ (%)")
      kable(age_summary_table_formatted,
            col.names = col_names,
            caption = paste("Age Distribution Summary by Municipality", if(latest_years$demographic > 0) paste("(", latest_years$demographic, ")") else "")) %>%
        kable_styling(bootstrap_options = c("striped", "hover"),
                      full_width = FALSE,
                      position = "center") %>%
        column_spec(1, bold = TRUE, color = "#0e1d6b") %>%
        row_spec(0, bold = TRUE, color = "white", background = "#007CC2")
    } else {
      cat("Unable to create age summary table due to data structure issues.")
    }
  } else {
    cat("No age category variables available for summary table.")
  }
} else {
  cat("No data available for age summary table.")
}
```

### 3. Healthcare Cost Analysis

```{r healthcare-costs, echo=FALSE, fig.width=10, fig.height=6}
# Create healthcare cost analysis
if (nrow(selected_data) > 0) {
  # Check which cost variables are available
  cost_vars <- c("ZVWKOSTENTOTAAL_MEAN", "ZVWKZIEKENHUIS_MEAN", "ZVWKHUISARTS_MEAN", "ZVWKFARMACIE_MEAN")
  available_cost_vars <- cost_vars[cost_vars %in% names(selected_data)]
  
  if (length(available_cost_vars) > 0) {
    # Prepare data for plotting
    cost_data <- selected_data %>%
      select(GMN, all_of(available_cost_vars)) %>%
      pivot_longer(cols = all_of(available_cost_vars), 
                   names_to = "Cost_Type", 
                   values_to = "Cost") %>%
      filter(!is.na(Cost))
    
    if (nrow(cost_data) > 0) {
      # Create cost labels
      cost_labels <- c(
        "ZVWKOSTENTOTAAL_MEAN" = "Total Healthcare Cost",
        "ZVWKZIEKENHUIS_MEAN" = "Secondary Care Cost",
        "ZVWKHUISARTS_MEAN" = "Primary Care Cost",
        "ZVWKFARMACIE_MEAN" = "Pharmacy Cost"
      )
      
      cost_plot <- cost_data %>%
        mutate(Cost_Type = cost_labels[Cost_Type]) %>%
        ggplot(aes(x = GMN, y = Cost, fill = Cost_Type)) +
        geom_col(position = "dodge") +
        labs(
          title = "Healthcare Costs by Municipality",
          x = "Municipality",
          y = "Average Cost (€)",
          fill = "Cost Type"
        ) +
        theme_minimal() +
        theme(
          plot.title = element_text(size = 18, face = "bold", color = "#0e1d6b"),
          axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "bottom"
        ) +
        scale_fill_brewer(palette = "Set1") +
        scale_y_continuous(labels = comma)
      
      print(cost_plot)
    } else {
      cat("No valid cost data available for plotting.")
    }
  } else {
    cat("No healthcare cost variables available in the dataset.")
  }
} else {
  cat("No data available for healthcare cost analysis.")
}
```

### 4. Primary Care Utilization

```{r primary-care, echo=FALSE, fig.width=10, fig.height=6}
# Create primary care utilization chart
if (nrow(selected_data) > 0 && "ZVWKHUISARTS_user" %in% names(selected_data)) {
  primary_care_data <- selected_data %>%
    select(GMN, ZVWKHUISARTS_user, Total_Population) %>%
    filter(!is.na(ZVWKHUISARTS_user) & !is.na(Total_Population)) %>%
    mutate(
      Users = ZVWKHUISARTS_user * Total_Population,
      Non_Users = Total_Population - Users,
      User_Percentage = ZVWKHUISARTS_user * 100
    )
  
  if (nrow(primary_care_data) > 0) {
    # Create percentage chart
    percentage_plot <- primary_care_data %>%
      ggplot(aes(x = reorder(GMN, User_Percentage), y = User_Percentage, fill = GMN)) +
      geom_col() +
      labs(
        title = "Primary Care Utilization by Municipality",
        x = "Municipality",
        y = "Percentage of Population Using Primary Care (%)"
      ) +
      theme_minimal() +
      theme(
        plot.title = element_text(size = 18, face = "bold", color = "#0e1d6b"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none"
      ) +
      scale_fill_brewer(palette = "Set1") +
      scale_y_continuous(labels = function(x) paste0(x, "%"))
    
    print(percentage_plot)
  } else {
    cat("No valid primary care data available for plotting.")
  }
} else {
  cat("Primary care data not available.")
}
```

### 5. Medication Usage Analysis

```{r medication-usage, echo=FALSE, fig.width=10, fig.height=6}
# Create medication usage analysis
if (nrow(selected_data) > 0 && "UniqueMed_Count5" %in% names(selected_data)) {
  medication_data <- selected_data %>%
    select(GMN, UniqueMed_Count5, Total_Population) %>%
    filter(!is.na(UniqueMed_Count5) & !is.na(Total_Population)) %>%
    mutate(
      Polypharmacy_Users = UniqueMed_Count5 * Total_Population,
      Non_Polypharmacy_Users = Total_Population - Polypharmacy_Users,
      Polypharmacy_Percentage = UniqueMed_Count5 * 100
    )
  
  if (nrow(medication_data) > 0) {
    # Create polypharmacy chart
    polypharmacy_plot <- medication_data %>%
      ggplot(aes(x = reorder(GMN, Polypharmacy_Percentage), y = Polypharmacy_Percentage, fill = GMN)) +
      geom_col() +
      labs(
        title = "Polypharmacy Usage by Municipality",
        subtitle = "Percentage of Population Using 5+ Medications",
        x = "Municipality",
        y = "Percentage of Population (%)"
      ) +
      theme_minimal() +
      theme(
        plot.title = element_text(size = 18, face = "bold", color = "#0e1d6b"),
        plot.subtitle = element_text(size = 14, color = "#666"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none"
      ) +
      scale_fill_brewer(palette = "Set1") +
      scale_y_continuous(labels = function(x) paste0(x, "%"))
    
    print(polypharmacy_plot)
  } else {
    cat("No valid medication data available for plotting.")
  }
} else {
  cat("Medication data not available.")
}
```

---

## Data Summary Tables

### Municipality Overview

```{r municipality-overview, echo=FALSE}
if (nrow(selected_data) > 0) {
  overview_table <- selected_data %>%
    select(GMN, Total_Population, AGE_MEAN, Income_MEAN) %>%
    mutate(
      Total_Population = format(Total_Population, big.mark = ","),
      AGE_MEAN = round(AGE_MEAN, 1),
      Income_MEAN = format(round(Income_MEAN), big.mark = ",")
    ) %>%
    rename(
      Municipality = GMN,
      Population = Total_Population,
      `Average Age` = AGE_MEAN,
      `Average Income (€)` = Income_MEAN
    )
  
  kable(overview_table, format = "html", align = c("l", "r", "r", "r")) %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
    column_spec(1, bold = TRUE)
} else {
  cat("No data available for municipality overview.")
}
```

### Healthcare Indicators Summary

```{r healthcare-summary, echo=FALSE}
if (nrow(selected_data) > 0) {
  # Check which healthcare variables are available
  healthcare_vars <- c("ZVWKOSTENTOTAAL_MEAN", "ZVWKZIEKENHUIS_MEAN", "ZVWKHUISARTS_MEAN", "ZVWKFARMACIE_MEAN", "ZVWKHUISARTS_user", "UniqueMed_Count5")
  available_healthcare_vars <- healthcare_vars[healthcare_vars %in% names(selected_data)]
  
  if (length(available_healthcare_vars) > 0) {
    healthcare_summary <- selected_data %>%
      select(GMN, all_of(available_healthcare_vars)) %>%
      mutate(across(all_of(available_healthcare_vars), ~round(., 2)))
    
    # Rename columns for better display
    colnames(healthcare_summary) <- c("Municipality", 
                                      "Total Healthcare Cost" = "ZVWKOSTENTOTAAL_MEAN",
                                      "Secondary Care Cost" = "ZVWKZIEKENHUIS_MEAN",
                                      "Primary Care Cost" = "ZVWKHUISARTS_MEAN",
                                      "Pharmacy Cost" = "ZVWKFARMACIE_MEAN",
                                      "Primary Care Users (%)" = "ZVWKHUISARTS_user",
                                      "Polypharmacy Users (%)" = "UniqueMed_Count5")[colnames(healthcare_summary)]
    
    kable(healthcare_summary, format = "html", align = c("l", rep("r", length(available_healthcare_vars)))) %>%
      kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
      column_spec(1, bold = TRUE)
  } else {
    cat("No healthcare indicators available in the dataset.")
  }
} else {
  cat("No data available for healthcare summary.")
}
```

---

## Data Quality and Limitations

<div style="background-color: #fff3cd; padding: 20px; border-radius: 8px; margin: 20px 0; border-left: 4px solid #ffc107;">
  <h3 style="color: #856404; margin-bottom: 15px;">⚠️ Data Quality Notes</h3>
  <ul style="color: #856404; line-height: 1.6;">
    <li><strong>Data Source:</strong> ELAN Dashboard - Municipality-level aggregated data</li>
    <li><strong>Geographic Coverage:</strong> ELAN Region Municipalities</li>
    <li><strong>Time Period:</strong> Available years vary by municipality and indicator</li>
    <li><strong>Missing Data:</strong> Some variables may have missing values for certain years or municipalities</li>
    <li><strong>Definitions:</strong> Variable definitions are based on ELAN data dictionary</li>
    <li><strong>Aggregation Level:</strong> Data is aggregated at the municipality level, not individual patient level</li>
  </ul>
</div>

---

## Contact Information

<div style="background-color: #e8f4fd; padding: 20px; border-radius: 8px; margin: 20px 0; border-left: 4px solid #007CC2;">
  <h3 style="color: #0e1d6b; margin-bottom: 15px;">📞 Contact Information</h3>
  <p style="color: #333; line-height: 1.6;">
    For questions about this report or the ELAN Dashboard, please contact:
  </p>
  <p style="color: #333; line-height: 1.6; margin-top: 10px;">
    <strong>ELAN DCC</strong><br>
    Email: <a href="mailto:ELAN.DCC@lumc.nl" style="color: #007CC2;">ELAN.DCC@lumc.nl</a><br>
    Website: <a href="https://www.lumc.nl/over-het-lumc/partners/elan/" target="_blank" style="color: #007CC2;">https://www.lumc.nl/over-het-lumc/partners/elan/</a>
  </p>
</div>

---

*Report generated on `r Sys.Date()` at `r Sys.time()`* 
*Report generated on `r Sys.Date()` at `r Sys.time()`* 